Calculating Biological Module Enrichment or Depletion and Visualizing Data on Large-scale Molecular Maps with the R packages \pkg{ACSNMineR} and \pkg{RNaviCell}

### 2016-02-10 jss/article.tex
[skip ci] update

We loaded the data as a matrix in R and calculated the enrichment in ACSN maps
with the ACSNMineR function \code{enrichment}. The results are displayed in
table \ref{tab:table_gbm}. There are 16 modules significantly enriched in the
Cell Cycle, DNA repair, EMT motility and Survival. Cell matrix adhesions and ECM
(extra cellular matrix), part of the EMT motility map, are the modules with
highest significance (corrected p-values of 2.55e-11 and 2.34e-9 respectively).

@@

\begin{tabular}{l|c|c|c}
\hline
module & Mod. size & Nb genes & Cor. pval\\ 
\hline
CellCycle:G2\_M\_CHECKPOINT & 14 & 8 & 3.92e-02\\
CellCycle:S\_PHASE\_CHECKPOINT & 6 & 5 & 3.92e-02\\
DNA\_repair:S\_PHASE\_CHECKPOINT & 45 & 19 & 1.58e-02\\
EMT\_motility & 621 & 176 & 1.06e-03\\
EMT\_motility:CELL\_MATRIX\_ADHESIONS & 72 & 44 & 2.55e-11\\
EMT\_motility:ECM & 137 & 64 & 2.34e-09\\
EMT\_motility:EMT\_REGULATORS & 615 & 173 & 1.49e-03\\
EMT\_motility:CYTOSKELETON\_POLARITY & 154 & 47 & 3.92e-02\\
EMT\_motility:DESMOSOMES & 29 & 15 & 4.62e-03\\
ACSN\_master:CELL\_MATRIX\_ADHESIONS & 72 & 44 & 2.55e-11\\
ACSN\_master:DESMOSOMES & 29 & 15 & 4.62e-03\\
ACSN\_master:S\_PHASE\_CHECKPOINT & 46 & 20 & 8.23e-03\\
ACSN\_master:ECM & 138 & 64 & 2.56e-09\\
ACSN\_master:EMT\_REGULATORS & 625 & 177 & 1.06e-03\\
ACSN\_master:CYTOSKELETON\_POLARITY & 154 & 47 & 3.92e-02\\
Survival:ECM & 23 & 11 & 3.92e-02\\
\hline
\end{tabular}
\end{table}

Visualization of the list of glioblastoma mutated genes is shown on figure
\ref{fig:gbm}. This figure was generated with the ACSNMineR commands detailed
below. Results of the enrichment test correlate well with the visualization on
the map, with a high density of low and high frequency mutated genes in the EMT
motility , survival, cell cycle and DNA repair regions (maps) of the global ACSN
map.

>>>>

We loaded the data as a matrix in R and calculated the enrichment in ACSN maps
with the ACSNMineR function \code{enrichment}. The results are displayed in
table \ref{tab:table_gbm}. There are 6 modules significantly enriched in the
DNA repair and EMT motility maps. Cell matrix adhesions and ECM
(extra cellular matrix), part of the EMT motility map, are the modules with
highest significance. The EMT motility map is significantly enriched at the
global map level (second line in the table).

@@

\begin{tabular}{l|c|c|c}
\hline
module & Mod. size & Nb genes & Cor. pval\\ 
\hline
DNA\_repair:S\_PHASE\_CHECKPOINT &   45 &      19 &     0.008\\
EMT\_motility  &  635  &   181  &   0.0002\\
EMT\_motility:CELL\_MATRIX\_ADHESIONS   &   73  &    45  &    3.73e-12\\
EMT\_motility:CYTOSKELETON\_POLARITY   &   154 &    47  &    0.022\\
EMT\_motility:DESMOSOMES & 29  &    15  &    0.002\\
EMT\_motility:ECM    &    147  &   69   &   9.77e-11\\
EMT\_motility:EMT\_REGULATORS  &   629   &  178  &   0.0002\\
\hline
\end{tabular}
\end{table}

Visualization of the list of glioblastoma mutated genes is shown on figure
\ref{fig:gbm}. This figure was generated with the ACSNMineR commands detailed
below. Results of the enrichment test correlate well with the visualization on
the map, with a high density of low and high frequency mutated genes in the EMT
motility and DNA repair regions (maps) of the global ACSN map. Although they
are not statistically significant, quite high densities can also be seen in
other regions of the map.
